Injury Outcomes

### 2023-08-17 injury_tables.Rmd
Updated injury output calculations

---
title: "Injrury Outcomes - WHW matrices"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---

@@

# NOTE: the following scenario order is assumed: 'bi_sc', 'car_sc', 'bus_sc','mc_sc','w_sc' and if there are less than 5 scenarios, the last scenarios can be missing but all of the scenarios in the middle need to exist. e.g. if you want the public transport and the walking scenario, you still need to create a car, motorcycling and bicycling scenario!

@@

```{js echo=FALSE}
window.location.href='#Who_Hit_Whom_matrix_(normalized_to_annual_figures)';
```

```{r loadLibraries, echo = F, message = F}

st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
# much better results. Your mileage may vary

@@

```{r load_objects = "asis"}
io <- readRDS("results/multi_city/io.rds")
# Assumes that multi_city_script.R has been run till 
# Get names of cities from the io object
cities <- names(io)[!names(io) %in% 'scen_prop']

@@

round_to <- 1

# Set plot_mmets to F
plot_mmets <- F

sum_and_round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}
round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

get_qualified_scen_name <- function(cs){
  qualified_scen_name <- ""
  if (cs == 'baseline')
    qualified_scen_name <- 'Baseline'
  else if(cs == "sc_walk")
    qualified_scen_name <- 'Walking'
  else if(cs == "sc_cycle")
    qualified_scen_name <- 'Bicycling'
  else if(cs == "sc_car")
    qualified_scen_name <- 'Car'
  else if(cs == "sc_motorcycle")
    qualified_scen_name <- 'Motorcycling'
  else if(cs == "sc_bus")
    qualified_scen_name <- 'Public Transport'

@@

```{r preprocessing}

collapse_ages <- function(data, ages = c('15-49', '50-69'), age_lab = 'age_cat'){
  target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
  target_max_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][2]))
  min_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][1]))
  max_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][2]))
  genders <- unique(data$sex)
  if (ncol(data) > 3) {
    reformatted <- do.call(rbind,lapply(1:length(ages),
                                        function(x) t(sapply(genders,
                                                             function(y) 
                                                               if(ncol(data)>3) colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                                             else sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                        ))
    ))
    data.frame(age_cat=rep(ages,each=2),sex=rep(genders,2),reformatted,stringsAsFactors=F)
  }else{
    reformatted <- do.call(c,lapply(1:length(ages),
                                    function(x) sapply(genders,
                                                       function(y) 
                                                         sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                    )
    ))
    data.frame(age=rep(ages,each=2),sex=rep(genders,2),population=as.numeric(reformatted),stringsAsFactors=F)
  }
}

scen_prop <- io$scen_prop
io <- io[-2]
# re-format some age and injury names
for (i in 1:length(io)) {
  io[[i]]$demographic$age[io[[i]]$demographic$age == '5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$deaths <- io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
}
for (city in cities)
  for (type in c('hb','pathway_hb'))
    for (out in c('deaths','ylls'))
      io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]]) # only keep two age categories

for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')  # only keep two age categories

pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population))) # find number of people within two age categories
pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths)=='scen1_deaths_inj') # find col number of injury data
ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
pa_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
scen_names <- rownames(scen_prop)

@@

# Who Hit Whom matrix (normalized to annual figures)

Who hit whom matrix for all scenarios for each case study

## Who hit whom matrix of each case study (by baseline + scenarios) {#whw}

```{r whw, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}

@@

for (city in cities) {

@@

  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){ # if upper and lower bounds are given, this is NA, ub, lb
    # pref <- "ub"
    pref_name <- ""
    if (!is.na(pref))
      pref_name <- paste0("_", pref) # if pref is lb or ub e.g.

@@

    for (cs in names(io[[city]]$outcomes$whw)){ # Baseline, Scenario 1, Scenario 2, etc.
      # cs <- 'baseline'

      if (is.null(io[[city]]$outcomes$whw[[cs]][[whw_qn]])){
        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # if only data for the given name exists for nov but not for whw, only create data frame from nov part
          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
          td2$mode <- 'NOV'

          td2 <- td2 %>% dplyr::select(mode, names(.))

          td3 <- td2
        }

@@

      cat('####', qualified_scen_name, ' ', ifelse(is.na(pref), "mean", pref), ' WHW', '\n')

@@

      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))] # remove column modes that do not appear in list of standardized modes

      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),] # remove rows with mode names that do not appear in list of standardized modes

      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))  # convert numbers to characters

@@

@@

      if (length(city_df[[cs]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) # combine mean, lb and ub counts into one table for each scenario
    }

  }
  # create df with location information
  cityname <- city
  if (city == cities[[1]]){
    location <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
  } else {
    dummy_row <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
    location <- rbind(location, dummy_row)
  }

  # create dataframe with original fatality counts adjusted by weight and injury_reporting_rate

  # Baseline - whw
  injury_data_count <- as.data.frame(io[[city]]$injury_table$whw)
  injury_data_count <- injury_data_count %>% rename(value = count, str_mode = strike_mode) # rename columns

@@

  city_baseline_count$city <- city
  city_baseline_count$scenario <- 'Baseline'
  city_baseline_count$country <- io[[city]]$location$country
  city_baseline_count$continent <- io[[city]]$location$continent

  if (city == cities[[1]]){
    baseline_counts <- city_baseline_count
  } else {
    baseline_counts <- rbind(baseline_counts, city_baseline_count)
  }

@@

  baseline_pred <- dummy_distances2 %>% filter(scenario == 'Baseline')
  baseline_pred$scenario <- 'Baseline_predicted'

@@

  ####### create duration df based on true distances and speed
  speed_city <- io[[city]]$vehicle_inventory %>% dplyr::select(-c( "PM_emission_inventory",  "CO2_emission_inventory"))
  speed_city$city <- cityname

@@

  # dummy_dur <- io[[city]]$dur
  # 
  # dummy_dur2 <- pivot_longer(dummy_dur, cols = -c(stage_mode), names_to = 'scenario', values_to = 'mode_duration')
  # 
  # baseline_pred <- dummy_dur2 %>% filter(scenario == 'Baseline')
  # baseline_pred$scenario <- 'Baseline_predicted'
  # 
  # dummy_dur3 <- rbind(dummy_dur2, baseline_pred)
  # 
  # dummy_dur3$city <- city
  # 
  # if (city == cities[1]){
  #   duration <- dummy_dur3
  # } else {
  #   duration <- rbind(duration,dummy_dur3)
  # }
}

# calculate duration
duration <- left_join(true_distances, speed_df, by = c('city','stage_mode' ))
duration$mode_duration <- duration$mode_distance / duration$speed
duration <- duration %>% dplyr::select(-c(mode_distance, speed))

@@

```{r whw_for_all_cs, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
st <- list()
file_list <- list()
td <- NULL
whw_list <- list()
for (cs in names(all_whw)){ # loop through scenarios
  # cs <- names(all_whw)[1]

  td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = c("mode", "measure")) %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))

  td <- (td[!duplicated(td), ])

  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total')) # re-arrange order of rows

@@

  colnames(td) = gsub("auto_rickshaw", "ar", colnames(td))

  whw_lng <- reshape2::melt(td) # create df with 1 column containing all the fatality counts

  col_split <- stringr::str_split(whw_lng$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column

  # tidy and rename some of the columns
  whw_lng <- cbind(whw_lng, col_split)
  names(whw_lng)[5] <- 'strike_mode'
  names(whw_lng)[6] <- 'city'
  whw_lng <- whw_lng %>% dplyr::select(-variable)
  whw_lng <- whw_lng %>% dplyr::filter(strike_mode != "rowSum")
  whw_lng <- whw_lng %>% rename(str_mode = mode, cas_mode = strike_mode)

  if (nrow(whw_lng[whw_lng$cas_mode == 'ar',]) > 0){
    whw_lng[whw_lng$cas_mode == 'ar',]$cas_mode <- "auto_rickshaw"
  }

  if (nrow(whw_lng[whw_lng$str_mode == 'ar',]) > 0){
    whw_lng[whw_lng$str_mode == 'ar',]$str_mode <- "auto_rickshaw"
  }

  whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode)) # assign factors

  qualified_scen_name <- 'Public Transport'
  scen <- 'Scenario'

@@

  whw_lng$scenario <- qualified_scen_name

  if (length(whw_list) == 0)
    whw_list <- whw_lng
  else
    whw_list <- rbind(whw_list, whw_lng)

  # f <- ggplot(data = whw_lng) +
  #   aes(x = str_mode, fill = cas_mode, weight = value) +
  #   geom_bar(position = "stack") +
  #   scale_fill_brewer(palette = "Dark2") +
  #   labs(x = "Strike Mode", y = "# of incidents", title = paste("WHW - ", qualified_scen_name, scen)) +
  #   coord_flip() +
  #   facet_wrap(vars(city), scales = "free_y")
  # 
  # file_list[[cs]] <- f

  #ggsave(f, file = paste0("results/multi_city/whw_matrices/plots/whw_", cs,".png"), limitsize = F, dpi = 300, scale = 1.5)

  d <- paste0("results/multi_city/whw_matrices/interactive_plots/whw_", cs,".html")

  # htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
  #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")

@@

# add country and continent information
whw_list <- left_join(whw_list, location, by = 'city')

# add original counts and rename Baseline to Baseline predicted
whw_list$scenario <- ifelse(whw_list$scenario == 'Baseline', 'Baseline_predicted',whw_list$scenario)

whw_list <- rbind(whw_list, baseline_counts)

# replace capital with lower letters
whw_list$str_mode <- tolower(whw_list$str_mode)

# update true distances scenarios
true_distances <- true_distances |> rowwise() |> mutate(scenario = get_qualified_scen_name(scenario))

# add mode distances
whw_list <- left_join(whw_list, true_distances, by = c('city', 'cas_mode','scenario'))

# combine active travel modes
active_travel <- whw_list %>% filter(cas_mode == 'cycle' | cas_mode == 'pedestrian')
active_travel <- active_travel %>% group_by(str_mode, measure, city, scenario, country, continent) %>% summarise(
                                  value = sum(value), mode_distance = sum(mode_distance), cas_mode ='active_travel')

# Remove total from whw long form, and write as csv
readr::write_csv(whw_list %>% filter(str_mode != 'Total'), paste0('results/multi_city/whw_matrices/whw_lng.csv')) # table containing all the fatality counts
```

## Who hit whom matrix for all case cities {#whw_all}

```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
### by casualty mode # This does contain NOV fatalities!

>>>>

---
title: "Injury Outcomes"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
---

@@

# Script to summarise injury outcomes

# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city, i.e. 
#          it assumes that the reference scenario is the Baseline scenario

#### It produces the following output documents:

## - html file containing the following information:

#   - fatality counts by strike and casualty mode for each scenario and city (incl lower and upper confidence interval limits were applicable), one table for each city, scenario and mean, upper or lower confidence interval values

#   - fatality counts giving lower, mean and upper bounds (where they exist) for each casualty mode by strike mode for each scenario and city, one table for each casualty mode, scenario and city

#   - tables showing the fatalities per billion km for each city, scenario and mean, upper and lower confidence interval values (where applicable).

#   - tables showing the number of deaths per 100,000 people for each city, scenario and mean, upper and lower confidence interval values (where applicable).

#   - tables showing the fatalities per 100 million hours for each city, scenario and mean, upper and lower confidence interval values (where applicable).

## - output .csv files containing the following information:

#   - The inj_counts.csv file ('results/multi_city/inj/inj_counts.csv') contains the fatality counts (mean and upper and lower confidence interval boundaries if required) for each city, scenario and casualty and strike mode pair. It also includes the observed baseline fatality counts and the active travel fatality counts (the sum of cycle and walk casualty mode fatalities)

# - The injury_risks_per_billion_kms.csv file ('results/multi_city/inj/injury_risks_per_billion_kms.csv') gives the fatalities per billion km travelled for each city, scenario and casualty mode. It also includes the upper and lower confidence interval boundary values if required and active travel counts

# - The injury_risks_per_100k_pop.csv file ('results/multi_city/inj/injury_risks_per_100k_pop.csv') gives the fatalities per 100k people for each city, scenario and casualty mode. It also includes the upper and lower confidence interval boundary values if required and active travel counts

# - The distances.csv file ('results/multi_city/inj/distances.csv') contains the population distances travelled per day for each city, scenario and mode

# - The injury_risks_per_100million_h.csv file ('results/multi_city/inj/injury_risks_per_100million_h.csv') gives the fatalities per 100 million hours travelled for each city, scenario and casualty mode. It also includes the upper and lower confidence interval boundary values if required and active travel counts

# - The inj_data.xlsx file (results/multi_city/inj/inj_data.xlsx) contains all injury output csv files

## The script performs the following steps, assuming that ITHIM-GLOBAL module has 
# been run and the ithim_objects object has been saved in "results/multi_city/io.rds":

# - read in ithim_objects and extract relevant information such as city and scenario names 
# - define various functions to summarise and print data and to re-name scenarios, also
#   define function to collapse the age categories into two categories (15-49 and 50-69)
# - calculate the health burden (yll and deaths) and the demographic information for those
#   two age categories (15-49 and 50-69)

# - loop through cities:
#   - loop through the mean injury values but also the upper and lower confidence boundary values if given:
#     - loop through scenarios incl baseline:
#       - create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
#         of nov or whw exists, only create dataframe with existing data), add row and column sums
#       - delete casualty and strike modes that do not exist in list of standardized modes
#       - print dataframe to html file
#       - create one list containing the predicted injury counts for each city, each scenario and for each scenario
#         both the mean and the upper and lower confidence interval limits are given if they exist
#       - create a dataframe with the original fatality counts for each casualty and strike mode pair
#         (adjusted by weight and injury reporting rate)
#       - create dataframe containing the mode distances for all scenarios including the predicted baseline
#       - create dataframes containing the population counts, mode speeds and durations travelled

# - create dataframe that contains the fatality counts for each city, scenario and casualty and strike mode pair.
#   It also includes the observed fatality counts and the active travel fatality counts (the sum of cycle and 
#   walk casualty mode fatalities) - ('results/multi_city/inj/inj_counts.csv') 
#   - print individual table for each city, scenario and mean, upper and lower confidence value fatality counts 
#     (where applicable) by casualty and strike mode to html file

# - For each city, scenario and casualty mode, print a table to the html file giving the fatality counts by strike mode
#   for the mean (and upper and lower confidence value boundaries) values

# - For each city and scenario give mean (and upper and lower confidence boundaries where required) fatalities per 
#   billion km travelled for each casualty and strike mode pair. Create html tables.

# - Create csv file ('injury_risks_per_billion_kms.csv') for fatalities per billion km travelled

# - Calculate the number of fatalities per 100,000 people for each city, scenario and mean, upper and lower confidence
#   interval values (where applicable). Print tables to .html file and also create a csv file ('injury_risks_per_100k_pop.csv')

# - Create dataframe containing all mode distances for each city and scenario ('distances.csv')

# - Calculate tables showing the fatalities per 100 million hours for each city, scenario and mean, upper
#   and lower confidence interval values (where applicable). Print tables to .html file

# - Create a csv file with the fatalities per 100 million hours (injury_risks_per_100million_h.csv)

# - Export all injury datasets into a single excel sheet (inj_data.xlsx)

@@

```{js echo=FALSE}
window.location.href='#Injury_outputs_(normalized_to_annual_figures)';
```

```{r loadLibraries, echo = F, message = F}
# set up 
st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives much better results. Your mileage may vary

@@

```{r load_objects = "asis"}
# Assumes that multi_city_script.R has been run - read in ithim_objects saved as io.rds
io <- readRDS("results/multi_city/io.rds")

# Get names of cities from the io object
cities <- names(io)[!names(io) %in% 'scen_prop']

@@

# define which decimal place to round to
round_to <- 1

# function to round and sum the data and then to print it 
# the html output document
sum_and_round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

# function to round and print data
round_and_print <- function(data,text=''){
  data <- lapply(data, function(x)round(x,round_to))
  for(city in cities) {
    print(kable(data[[city]], caption = paste(text, city)))
    cat("\n")
  }
}

# function to rename scenarios
get_qualified_scen_name <- function(cs){
  qualified_scen_name <- ""
  if (cs == 'baseline')
    qualified_scen_name <- 'Baseline predicted'
  else if(cs == "sc_walk")
    qualified_scen_name <- 'Walking'
  else if(cs == "sc_cycle")
    qualified_scen_name <- 'Bicycling'
  else if(cs == "sc_car")
    qualified_scen_name <- 'Car'
  else if(cs == "sc_motorcycle")
    qualified_scen_name <- 'Motorcycling'
  else if(cs == "sc_bus")
    qualified_scen_name <- 'Public Transport'
  else if(cs == "Baseline predicted")
    qualified_scen_name <- 'Baseline predicted'

@@

```{r preprocessing}
# start the processing

# function to collapse the given age categories into only two categories
# '15-49' and '50-69'
collapse_ages <- function(data, ages = c('15-49', '50-69'), age_lab = 'age_cat'){
  target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
  target_max_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][2]))
  min_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][1]))
  max_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][2]))
  genders <- unique(data$sex)
  if (ncol(data) > 3) {
    reformatted <- do.call(rbind,lapply(1:length(ages),
                                        function(x) t(sapply(genders, function(y) 
                                          if(ncol(data)>3)     colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                                             else sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                        ))
    ))
    data.frame(age_cat=rep(ages,each=2),sex=rep(genders,2),reformatted,stringsAsFactors=F)
  }else{
    reformatted <- do.call(c,lapply(1:length(ages),
                                    function(x) sapply(genders,
                                                       function(y) 
                                                         sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
                                    )
    ))
    data.frame(age=rep(ages,each=2),sex=rep(genders,2),population=as.numeric(reformatted),stringsAsFactors=F)
  }
}

scen_prop <- io$scen_prop
io <- io[-2]

# re-format some age and injury names
for (i in 1:length(io)) {
  io[[i]]$demographic$age[io[[i]]$demographic$age == '5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
  io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$deaths <-      io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
}

# calculate the demographics for the two age categories '15-49' and '50-69'
for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')  # only keep two age categories

@@

# Injury outcome (normalized to annual figures)

Injury outcome for all scenarios for each scenario including the predicted and observed baseline figures

## Injury outcome for each scenario (incl predicted and observed baseline figures)

```{r whw, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
# create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
# of nov or whw exists, only create dataframe with existing data), add row and column sums
# create distance, speed and duration data frames

@@

for (city in cities) { # loop through all cities

@@

  # loop through either mean values only or, if upper and lower confidence intervals are given, loop
  # through them as well
  #  if upper and lower bounds are given, 'pref' is set to 'NA', 'ub' and 'lb'
  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){ 
    # pref <- "ub"
    pref_name <- ""
    if (!is.na(pref))
      pref_name <- paste0("_", pref) # if pref is lb or ub e.g.

@@

    # loop through scenarios incl baseline
    for (cs in names(io[[city]]$outcomes$whw)){ 
      # cs <- 'baseline'

      # if data only exists for nov but not for whw, only create data frame from nov part
      if (is.null(io[[city]]$outcomes$whw[[cs]][[whw_qn]])){ # if no 'whw' part exists
        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # but if 'nov' part exists

          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
          td2$mode <- 'NOV'
          td2 <- td2 %>% dplyr::select(mode, names(.))
          td3 <- td2
        }

@@

      cat('####', qualified_scen_name, ' ', ifelse(is.na(pref), "mean", pref), ' fatalities by cas and str mode', '\n')

@@

      # remove column modes that do not appear in list of standardized modes
      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))]

      # remove rows with mode names that do not appear in list of standardized modes
      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),]

      # convert numbers to characters
      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))
                                ) %>%  mutate_if(is.numeric, ~as.character(.))

@@

      cat(knit_print(ft)) # print table to html file

@@

      if (length(city_df[[cs]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) # combine mean, lb and ub counts into one table for each scenario
    }
  }

  # create df with location information
  cityname <- city
  if (city == cities[[1]]){
    location <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
  } else {
    dummy_row <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
    location <- rbind(location, dummy_row)
  }

  # create dataframe with original fatality counts adjusted by weight and injury_reporting_rate 
  # Baseline - whw
  injury_data_count <- as.data.frame(io[[city]]$injury_table$whw)
  injury_data_count <- injury_data_count %>% rename(value = count, str_mode = strike_mode) # rename columns

@@

  # add further city, scenario and location information
  city_baseline_count$city <- city
  city_baseline_count$scenario <- 'Baseline'
  city_baseline_count$country <- io[[city]]$location$country
  city_baseline_count$continent <- io[[city]]$location$continent

  # create one dataframe for all cities
  if (city == cities[[1]]){
    baseline_counts <- city_baseline_count
  } else {
    baseline_counts <- rbind(baseline_counts, city_baseline_count)
  }

@@

  baseline_pred <- dummy_distances2 %>% filter(scenario == 'baseline' | scenario == 'Baseline')
  baseline_pred$scenario <- 'Baseline'

  # update scenario names
  dummy_distances2 <- dummy_distances2 |> rowwise() |> mutate(scenario = get_qualified_scen_name(scenario))

@@

  # create speed dataframe
  speed_city <- io[[city]]$vehicle_inventory %>% dplyr::select(-c( "PM_emission_inventory",  "CO2_emission_inventory"))
  speed_city$city <- cityname

@@

# calculate duration based on distance and speed
duration <- left_join(true_distances, speed_df, by = c('city','stage_mode' ))
duration$mode_duration <- duration$mode_distance / duration$speed
duration <- duration %>% dplyr::select(-c(mode_distance, speed))

@@

```{r whw_for_all_cs, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}

# create one dataframe for all cities and scenarios (incl original baseline counts) containing one column with fatality counts (and columns for city, strike mode, cas mode, scenario, distances, location)
# calculate active_travel fatality counts (where cas mode is either walk or cycle) and add to fatality dataframe
# write csv file

st <- list()
file_list <- list()
td <- NULL
inj_counts_list <- list()
for (cs in names(all_whw)){ # loop through scenarios
  # cs <- names(all_whw)[1]

  # join list elements of strike and cas mode fatilty predictions for all cities for the given scenario
  td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = c("mode", "measure")) %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))

  td[is.na(td)] <- 0 # set na values to 0

  # remove duplicate values
  td <- (td[!duplicated(td), ])

  # re-arrange order of rows
  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total'))

@@

  # re-name auto_rickshaw as ar
  colnames(td) = gsub("auto_rickshaw", "ar", colnames(td))

  # create df with 1 column containing all the fatality counts
  inj_counts <- reshape2::melt(td)

  col_split <- stringr::str_split(inj_counts$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column

  # tidy and rename some of the columns
  inj_counts <- cbind(inj_counts, col_split)
  names(inj_counts)[5] <- 'strike_mode'
  names(inj_counts)[6] <- 'city'
  inj_counts <- inj_counts %>% dplyr::select(-variable)
  inj_counts <- inj_counts %>% dplyr::filter(strike_mode != "rowSum")
  inj_counts <- inj_counts %>% rename(str_mode = mode, cas_mode = strike_mode)

  # re-name 'ar' to auto_rickshaw again ('_' in 'auto_rickshaw' would have caused issues when splitting the city and strike mode column)
  if (nrow(inj_counts[inj_counts$cas_mode == 'ar',]) > 0){
    inj_counts[inj_counts$cas_mode == 'ar',]$cas_mode <- "auto_rickshaw"
  }

  if (nrow(inj_counts[inj_counts$str_mode == 'ar',]) > 0){
    inj_counts[inj_counts$str_mode == 'ar',]$str_mode <- "auto_rickshaw"
  }

  inj_counts$str_mode <- factor(inj_counts$str_mode, levels = unique(inj_counts$str_mode)) # assign factors

  #qualified_scen_name <- 'Public Transport'
  scen <- 'Scenario'

@@

  inj_counts$scenario <- qualified_scen_name

  if (length(inj_counts_list) == 0)
    inj_counts_list <- inj_counts
  else
    inj_counts_list <- rbind(inj_counts_list, inj_counts)

@@

# add country and continent information
inj_counts_list <- left_join(inj_counts_list, location, by = 'city')

# add original counts and rename Baseline to Baseline predicted
#inj_counts_list$scenario <- ifelse(inj_counts_list$scenario == 'Baseline', 'Baseline_predicted',inj_counts_list$scenario)

inj_counts_list <- rbind(inj_counts_list, baseline_counts)

# replace capital with lower letters
inj_counts_list$str_mode <- tolower(inj_counts_list$str_mode)

# add mode distances
inj_counts_list <- left_join(inj_counts_list, true_distances, by = c('city', 'cas_mode','scenario'))

# combine active travel modes
active_travel <- inj_counts_list %>% filter(cas_mode == 'cycle' | cas_mode == 'pedestrian')
active_travel <- active_travel %>% group_by(str_mode, measure, city, scenario, country, continent) %>% summarise(
                                  value = sum(value), mode_distance = sum(mode_distance), cas_mode ='active_travel')

# add active travel modes to inj_counts_list
inj_counts_list <- rbind(inj_counts_list, active_travel)

# add totals for each cas_mode
inj_counts_list <- inj_counts_list %>% filter(str_mode != 'Total' & str_mode != 'total')
inj_counts_list_total <- inj_counts_list %>% group_by(measure, cas_mode, city, scenario, country, continent, mode_distance) %>%
                                        summarise(str_mode = 'Total', value = sum(value))
inj_counts_list <- rbind(inj_counts_list, inj_counts_list_total)
